### 2014-03-14 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago Fire: rename directory, update report and python

\begin{table}[!h]
\centering
\captionof{table}{Fire Event Timeline}\label{tab:fire_info}
\begin{tabularx}{0.8\textwidth}{c p{3cm}}
\toprule[1.5pt]
Incident Time & Fire Behavior / Fireground Operation  \\
\midrule
17:15    & Reported working fire with heavy smoke coming from the front and rear of the structure's attic. \\
17:16    & Dispatch for a {\bf Still Alarm} \\
17:20    & Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear. \\  
17:21:00 & E123 Capt. and FF on 2$^{nd}$ floor with 1 3/4-in hose line. \\
17:21:30 & 1$^{st}$ floor exterior porch door opened. \\
17:23    & Heavy fire report in rear stairwell and enclosed porch. No response from E123: Capt. or FF. \\
17:24    & Water applied to attic from the rear exterior position. \\
17:27    & IC calls a Mayday \\
17:29    & E123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabularx}
\end{table}

\chapter{Fire Dynamics Simulator and Smokeview}
FDS is a computational fluid dynamics (CFD) model that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires.  Within a CFD model, the room or building is divided into small three-dimensional rectangular control volumes or computational cells.  The cells are contained together within one larger volume known as a computational domain.  The CFD model computes the density, velocity, temperature, pressure and species concentration of the gas in each cell.  Based on the laws of conservation of mass, momentum,  and energy, the model tracks the generation and movement of fire gasses. One of the most important advantages of FDS is that it is  mathematically verified \cite{FDS_Verification_Guide} and validated against fire test data to ensure that it provides the expected results, given sufficient input data \cite{FDS_Validation_Guide}.  A complete description of the FDS model is given in reference \cite{FDS_Math_Guide}.

Smokeview is a software tool designed to visualize the results of the FDS \cite{Smokeview_Users_Guide}.  Smokeview visualizes smoke and other attributes of the modeled fire using traditional scientific methods such as displaying tracer particle flow, two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data such as temperature and flow vectors showing flow direction and magnitude.  Smokeview allows the fire and fire movement to be visualized.  This is done by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS.  Smokeview also visualizes static data at particular times using 2D or 3D contours of data such as temperature and flow vectors showing flow direction and magnitude.

@@

The development of the model geometry and development of the input file is straight forward with guidance from the FDS User's Guide  \cite{FDS_Users_Guide} or third party software.  Engineering judgment, graduate level knowledge of fire dynamics, and modeling experience are required to  translate the information collected from the scene into appropriate and realistic computational information.  Examples of these requirements include:
\begin{itemize}
\item the computational cell size
\item model domain and boundaries conditions
\item the heat release rate (HRR) of the ignition source/initial fire
\item physical and thermal properties of fuels such as interior finish and furnishings
\end{itemize}
Since there is no all-inclusive, universal fire modeling input data base this type of information must be gathered from literature or generated for a given incident. For the simulations performed for this study, Table ~\ref{tab:mod_param} includes the relevant model input parameters. The sections below describe these inputs.

@@

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential structure with balloon frame construction and a wood gable roofing system.  The structure had a 17.15 m (56 ft) by 6.15 m (20 ft) footprint. This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a main living area of approximately 110 m$^2$ (2200 ft$^2$). Interior walls were constructed from wood frame dimensioned lumber with 1.27 cm (0.5 in) gypsum board over top. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels, one rear stairwell that connected the first and second levels and a second rear set of stairs which provided access to the attic. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the alpha side of the structure. Note that the majority of the damage is to the structure's attic.

@@

To estimate the fuel, fire size, and structure ventilation, on-scene videos and post-incident reports are used. Post-incident, the fire investigation concluded that the fire originated in the structure's attic \cite{NIOSH:Bowyer}. Post-fire pictures revealed extensive damage to the attic and second floor rear porch (cf. Fig.~(\ref{fig:charlie_ex})). Based on these findings, the FDS model includes two source fire locations: an initial fire in the attic and a secondary fire on the second floor porch landing. The attic and rear porch consist of a fuel load that is mainly wood: the rear porch was enclosed with wood planks exteriorly covered by foam board and vinyl siding, the roof was asphalt shingles over wood planking, and the attic and porch had oriented strand board (OSB) floors.

@@

If we represent wood with the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, and use carbon monoxide ($y_{\mathrm{CO}}=0.004 \; {\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$) and soot ($y_{\mathrm{C}}=0.015 \; {\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$) yields from the SFPE Handbook \cite{SFPE:Tewarson}, we can write a balanced chemical reaction for wood oxidation:

@@

The heat of combustion of wood used in this simulation is 16400 kJ/kg (7050 BTU/lb), based on data provided in the SFPE Handbook \cite{SFPE:Tewarson}. The heat of combustion quantifies the energy release per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

@@

Note that using the input lines above will invoke the simple chemistry reaction balance in FDS. This means that fuel and air react to form only CO$_2$, CO, H$_2$O, Soot, and N$_2$. If there is a desire for any other combustion products then the user must explicitly define those species and the chemical reaction that produces them \cite{FDS_Users_Guide}. Based on above input lines FDS uses the default, mixing-controlled fast chemistry combustion model. This mechanism  states that the rate of fuel consumption is proportional to both the local limiting reactant concentration and the local rate of mixing and extinction is based on the critical flame temperature approach \cite{FDS_Math_Guide}. While FDS provides users to the option to use a more complex finite-rate combustion mechanism, we do not have sufficient evidence to justify deviating from the default specifications.

To estimate the energy release per unit time and unit area (kW/m$^2$) of wood, we use work by Babrauskas and Grayson \cite{babrauskas1990}. Experiments were conducted in a cone calorimeter to determine the 5 minute average of HRRPUA (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Here, we take a conservative approach and assume the burning wood in the structure had the lowest energy density of the tested samples \cite{babrauskas1990}, similar to that of Douglas Fir plywood. Based on the range of heat flux exposures, the data was fit to determine HRRPUA ($y$) as a function of exposure ($x$):

@@

In Eq.~\ref{eq:heat_flux}, $a$ and $b$ are experimental constants. For Douglas Fir plywood these values were found to be 1.43 and 22 respectively. Typically radiative feedback from flaming combustion is between 20 kW/m$^2$ and 35 kW/m$^2$ \cite{tsaidrysdale}. If we estimate that the average radiation feedback to the pyrolyzing fuel is 20 kW/m$^2$ because of under-ventilated conditions within the structure, then following Eq.~\ref{eq:heat_flux} the average heat release per unit area (HRRPUA) for wood is 51 kW/m$^2$.

To calculate the total fire size, we need to determine the appropriate burnable area (exposed wood area) of the attic and second floor porch enclosure. Based on the floor plan of the structure, the underside of the roof is approximately 134 m$^2$ plus there is 105 m$^2$ of floor area in the attic. Examination of post incident images showed that the floor immediately over the back porch burned away. That area will be considered as part of the porch fire. Based on post-incident damage, we consider approximately 50 \% of the remaining floor area as fuel to the attic fire. To estimate the total attic fire size, we include the underside roof area (134 m$^2$) and the floor area (43 m$^2$). Using an HRRPUA of wood of 51 kW/m$^2$, results in a fire size of 9 MW. To estimate the porch fire size, we consider the the four walls of porch enclosure (42 m$^2$) and the floor and ceiling of the porch (24 m$^2$). The total area is 64 m$^2$. Using the same HRRPUA of 51 kW/m$^2$, the fire size on the porch is approximately 3.3 MW. The total fire size for this structure is estimated to be 12.3 MW. Since it is unknown when the fire started relative to notification of a fire or how fast the fire spread, several simplifications are made. First is that the attic fire starts at the beginning of the simulation and ramps up to 9 MW in 10 seconds. The porch fire starts 40 seconds into the simulation and ramps to 3.3 MW in 10 seconds. A ten second ramp to peak was used to prevent sharp velocity changes at the fire boundary condition in FDS. The second is that the sources of the gas phase wood fuel are of fixed area and at fixed locations (2 sources for attic, 1 source for porch). The following lines define the source fires for an FDS input file:

@@

\section{Materials}
\label{matl}
While the fire was simplified to constant area fuel sources (Section~\ref{fire}), defining the material properties of the walls and ceiling are still important from a heat transfer perspective. In this study, we define gypsum \cite{Quintiere:2} for the finished walls and ceilings on the main two floors of the structure and wood \cite{Incropera:1} for the unfinished portions of the attic and rear porch. The text provided below are the lines necessary to define these materials in an FDS input file.

@@

\section{Vents}
\label{Vents}
This simulation includes vents that account for the changes in ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). There are two ventilation scenarios that are considered in this study. The first is the baseline case where the ventilation times and ventilation areas represent our best understanding of the incident. The second case does not include the firefighter opening the rear door to make entry to the structure and the subsequent beta side wall breach. The times of the ventilation changes are provided in Table \ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires \cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of structure of this type. Therefore, leakage would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

\bigskip
\begin{minipage}{.8\linewidth}
\centering
\captionof{table}{Simulation Ventilation Timelines}\label{tab:vents}
\begin{tabular}{ccccc}
\toprule[1.5pt]
Baseline & No Rear Door Open & Ventilation Area [m$^2$] & Type & Structure Side \\
\midrule
0:00    &  0:00  & 0.11 & Window      & Charlie \\
0:40    &  0:40  & 0.08 & Breach (x2) & Alpha and Charlie \\
0:45    &  0:45  & 0.12 & Breach      & Beta \\
0:55    &  0:55  & 0.12 & Breach      & Charlie \\     
2:10    &  N/A   & 1.41 & Door        & Charlie \\
2:15    &  N/A   & 0.2  & Breach      & Beta \\        
2:40    &  2:40  & 0.80 & Door        & Interior \\
5:00    &  5:00  & \multicolumn{3}{c}{Simulation Ends}  \\
\bottomrule[1.25pt]
\end{tabular}\par
\bigskip
\end{minipage}

@@

For the simulation in this study, we can estimate a measure of how well the flow field is resolved by the non-dimensional expression $D^*/\dx$. Here $D^*$ is a characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide \cite{FDS_Users_Guide}, ``the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$.'' Here, $D$ is the physical diameter of the fire as input to the model. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the characteristic fire diameter to cell size ratio based on the 3.3 MW porch fire is 15.5 and for the 9 MW fire is 23. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 16 \cite{NUREG_1824}. The grid resolution used in the model falls within typical, engineering $D^*/\dx$ values. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10 cm. Fig.~(\ref{fig:geom_grid}) shows the 
alpha side (front) of the structure rendered in Smokeview with the 10 cm mesh.

@@

As discussed Section~\ref{geom}, the structure has 17.15 m by 6.15 m footprint with a total height of 8.56 m. To ensure adequately resolved fluid flow into and out of the structure the computational domain is extended beyond the volume of the structure. The total volume of the computational domain is 24 m by 10 m by 10 m. At 10 cm resolution this would result in 2.4 million computational cells. To make the simulation time more tractable, the domain was divided into 16 equal 150 thousand cell meshes to be processed in parallel. Fig.~(\ref{fig:mult_mesh}) shows the entire structure within the computational domain.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_mesh}
\caption{Alpha-Delta side of the structure within computational domain. Each of the 16 meshes is 3 m by 10 m by 5 m.}
\label{fig:mult_mesh}
\end{figure}

\chapter{Model Results}

@@

In Fig.~(\ref{fig:hrr}), the solid line is the FDS HRR from the venting conditions that represent our attempt to model the on-scene conditions. The dashed line represents the case where the rear door opening and subsequent beta side breach do not occur. The dash-dotted line is the prescribed FDS input HRR. We can see that within the first 10 seconds, there are under-ventilated conditions within the structure as the modeled HRR is below the prescribed HRR. The HRR first peaks at just under 9 MW at approximately 20 seconds into the simulation. The HRR proceeds to drop to 5 MW as the available oxygen inside the structure gets consumed. There is insufficient available oxygen in the structure to combust all of the fuel and combustion occurs at then vents. Fig.~(\ref{fig:smv_ext_fire}) illustrates exterior flaming combustion through the failed Charlie side window and attic breach just after the second fire ignites at 40s. Despite the fire becoming under-ventilated, the mass flow rate of fuel into the structure remains constant. As a result there is a buildup of pressure and excess fuel within the attic and back porch. After additional wall breaches at 45 and 55 seconds (see Table~\ref{tab:vents}), the HRR spikes to 9 MW before again dropping to about 6 MW. This peak occurs because the additional vents provide pathways for pressure relief. Unburned fuel at above ambient pressure flow out of the structure, mix with oxygen, and react. As the pressure drops, the flow of fuel out of the structure slows and the HRR drops.

From Fig.~(\ref{fig:hrr}), we see that the HRR for both FDS simulations are identical until 130 seconds (HRR $\approx$ 7 MW), when the rear door is opened in the the baseline (solid line) case. The opening of the rear door coincides with a breach on the beta side of the structure. 
The open door and breach act as additional pressure releases for the structure. The high pressure unburned fuel inside the structure flows toward ambient pressure of the external environment. The fuel mixes with oxygen, burns, and the HRR spikes from about 7 MW to 9.7 MW at 139 seconds. As the pressure differential drops, the HRR correspondingly drops toward 9 MW at 160 seconds. In the case where the rear door and beta wall penetration do not occur, the HRR steadily climbs from 7 MW at 130 seconds to 9 MW at 160 seconds.
At the time of door failure, the HRR in both cases was approximately 9 MW. While there is more total energy released in the case where the back door was opened, at the

@@

>>>>

\bigskip 
\begin{minipage}{.8\linewidth}
\centering
\captionof{table}{Fire Event Timeline}\label{tab:fire_info}
\begin{tabular}{c m{0.25in}}
\toprule[1.5pt]
Incident Time & Fire Behavior / Fireground Operation  \\
\midrule
17:15    & Reported working fire with heavy smoke coming from the front and rear of the structure's attic. \\
17:16    & Dispatch for a Still Alarm \\
17:20    & Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear. \\  
17:21:00 & E123 Capt. and FF on 2$^{nd}$ floor with 1 3/4-in hose line. \\
17:21:30 & 1$^{st}$ floor exterior porch door opened. \\
17:23    & Heavy fire report in rear stairwell and enclosed porch. No response from E123: Capt. or FF. \\
17:24    & Water applied to attic from the rear exterior position. \\
17:27    & IC calls a Mayday \\
17:29    & E123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\bigskip
\end{minipage}

\chapter{Fire Dynamics Simulator and Smokeview}
FDS is a computational fluid dynamics (CFD) model that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires.  Within a CFD model, the room or building is divided into small three-dimensional rectangular control volumes or computational cells.  The cells are contained together within one larger volume known as a computational domain.  The CFD model computes the density, velocity, temperature, pressure and species concentration of the gas in each cell.  Based on the laws of conservation of mass, momentum,  and energy, the model tracks the generation and movement of fire gasses. One of the most important advantages of FDS is that it is  mathematically verified~\cite{FDS_Verification_Guide} and validated against fire test data to ensure that it provides the expected results, given sufficient input data~\cite{FDS_Validation_Guide}.  A complete description of the FDS model is given in reference~\cite{FDS_Math_Guide}.

Smokeview is a software tool designed to visualize the results of the FDS~\cite{Smokeview_Users_Guide}.  Smokeview visualizes smoke and other attributes of the modeled fire using traditional scientific methods such as displaying tracer particle flow, two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data such as temperature and flow vectors showing flow direction and magnitude.  Smokeview allows the fire and fire movement to be visualized.  This is done by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS.  Smokeview also visualizes static data at particular times using 2D or 3D contours of data such as temperature and flow vectors showing flow direction and magnitude.

@@

The development of the model geometry and development of the input file is straight forward with guidance from the FDS User's Guide ~\cite{FDS_Users_Guide} or third party software.  Engineering judgment, graduate level knowledge of fire dynamics, and modeling experience are required to  translate the information collected from the scene into appropriate and realistic computational information.  Examples of these requirements include:
\begin{itemize}
\item the computational cell size
\item model domain and boundaries conditions
\item the heat release rate (HRR) of the ignition source/initial fire
\item physical and thermal properties of fuels such as interior finish and furnishings
\end{itemize}
Since there is no all-inclusive, universal fire modeling input data base this type of information must be gathered from literature or generated for a given incident. For the simulations performed for this study, Table ~\ref{tab:mod_param} includes the relevant model input parameters. The sections below describe these inputs.

@@

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential structure with balloon frame construction and a wood gable roofing system.  The structure had a 17.15 m (56 ft) by 6.15 m (20 ft) footprint (cf. Fig.~(\ref{fig:simp_geom})). This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a main living area of approximately 110 m$^2$ (2200 ft$^2$). Interior walls were constructed from wood frame dimensioned lumber with 1.27 cm (0.5 in) gypsum board over top. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels, one rear stairwell that connected the first and second levels and a second rear set of stairs which provided access to the attic. Entry to the structure can be made either in the front (Alpha side) or rear (Charlie side). The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch (cf. Fig.~(\ref{fig:simp_geom})). A second floor interior door, connecting the rear porch and the hallway, failed and rapidly changed the tenability conditions for firefighters present on the second floor. More discussion on interior conditions can be found in Chapter~\ref{results}. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the alpha side of the structure. Note that the majority of the damage is to the structure's attic.

\begin{figure}[h!]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[width=6in]{../Figures/Simple_1_metric} \\
\includegraphics[width=6in]{../Figures/Simple_2_metric}
\end{tabular*}
\caption{Top view of 1st (top) and 2nd (bottom) floor of structure. Stairs and key rooms are identified.}
\label{fig:simp_geom}
\end{figure}

@@

To estimate the fuel, fire size, and structure ventilation, on-scene videos and post-incident reports are used. Post-incident, the fire investigation concluded that the fire originated in the structure's attic~\cite{NIOSH:Bowyer}. Post-fire pictures revealed extensive damage to the attic and second floor rear porch (cf. Fig.~(\ref{fig:charlie_ex})). Based on these findings, the FDS model includes two source fire locations: an initial fire in the attic and a secondary fire on the second floor porch landing. The attic and rear porch consist of a fuel load that is mainly wood: the rear porch was enclosed with wood planks exteriorly covered by foam board and vinyl siding, the roof was asphalt shingles over wood planking, and the attic and porch had oriented strand board (OSB) floors.

@@

If we represent wood with the following chemical formula, $\C\H_{1.7}\O_{0.74}\N_{0.002}$, and use carbon monoxide ($y_{\mathrm{CO}}=0.004 \; {\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$) and soot ($y_{\mathrm{C}}=0.015 \; {\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$) yields from the SFPE Handbook~\cite{SFPE:Tewarson}, we can write a balanced chemical reaction for wood oxidation:

@@

The heat of combustion of wood used in this simulation is 16400 kJ/kg (7050 BTU/lb), based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the energy release per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

@@

Note that using the input lines above will invoke the simple chemistry reaction balance in FDS. This means that fuel and air react to form only CO$_2$, CO, H$_2$O, Soot, and N$_2$. If there is a desire for any other combustion products then the user must explicitly define those species and the chemical reaction that produces them~\cite{FDS_Users_Guide}. Based on above input lines FDS uses the default, mixing-controlled fast chemistry combustion model. This mechanism  states that the rate of fuel consumption is proportional to both the local limiting reactant concentration and the local rate of mixing and extinction is based on the critical flame temperature approach~\cite{FDS_Math_Guide}. While FDS provides users to the option to use a more complex finite-rate combustion mechanism, we do not have sufficient evidence to justify deviating from the default specifications.

To estimate the energy release per unit time and unit area (kW/m$^2$) of wood, we use work by Babrauskas and Grayson~\cite{babrauskas1990}. Experiments were conducted in a cone calorimeter to determine the 5 minute average of HRRPUA (kW/m$^2$) for several different types of wood over a range of radiant heat fluxes (kW/m$^2$). Here, we take a conservative approach and assume the burning wood in the structure had the lowest energy density of the tested samples~\cite{babrauskas1990}, similar to that of Douglas Fir plywood. Based on the range of heat flux exposures, the data was fit to determine HRRPUA ($y$) as a function of exposure ($x$):

@@

In Eq.~\ref{eq:heat_flux}, $a$ and $b$ are experimental constants. For Douglas Fir plywood these values were found to be 1.43 and 22 respectively. Typically radiative feedback from flaming combustion is between 20 kW/m$^2$ and 35 kW/m$^2$~\cite{tsaidrysdale}. If we estimate that the average radiation feedback to the pyrolyzing fuel is 20 kW/m$^2$ because of under-ventilated conditions within the structure, then following Eq.~\ref{eq:heat_flux} the average heat release per unit area (HRRPUA) for wood is 51 kW/m$^2$.

To calculate the total fire size, we need to determine the appropriate burnable area (exposed wood area) of the attic and second floor porch enclosure. Based on the floor plan of the structure, the underside of the roof is approximately 134 m$^2$ (1453 ft$^2$) plus there is 105 m$^2$ (1130 ft$^2$) of floor area in the attic. Examination of post incident images showed that the floor immediately over the back porch burned away. That area will be considered as part of the porch fire. Based on post-incident damage, we consider approximately 50 \% of the remaining floor area as fuel to the attic fire. To estimate the total attic fire size, we include the underside roof area (134 m$^2$) and the floor area (43 m$^2$ (463 ft$^2$)). Using an HRRPUA of wood of 51 kW/m$^2$, results in a fire size of 9 MW. To estimate the porch fire size, we consider the the four walls of porch enclosure (42 m$^2$) and the floor and ceiling of the porch (24 m$^2$ (258 ft$^2$)). The total area is 64 m$^2$ (689 ft$^2$). Using the same HRRPUA of 51 kW/m$^2$, the fire size on the porch is approximately 3.3 MW. The total fire size for this structure is estimated to be 12.3 MW. Since it is unknown when the fire started relative to notification of a fire or how fast the fire spread, several simplifications are made. First is that the attic fire starts at the beginning of the simulation and ramps up to 9 MW in 10 seconds. The porch fire starts 40 seconds into the simulation and ramps to 3.3 MW in 10 seconds. A ten second ramp to peak was used to prevent sharp velocity changes at the fire boundary condition in FDS. The second is that the sources of the gas phase wood fuel are of fixed area and at fixed locations (2 sources for attic, 1 source for porch). The following lines define the source fires for an FDS input file:

@@

\section{Materials}
\label{matl}
While the fire was simplified to constant area fuel sources (Section~\ref{fire}), defining the material properties of the walls and ceiling are still important from a heat transfer perspective. In this study, we define gypsum~\cite{Quintiere:2} for the finished walls and ceilings on the main two floors of the structure and wood~\cite{Incropera:1} for the unfinished portions of the attic and rear porch. The text provided below are the lines necessary to define these materials in an FDS input file.

@@

\section{Vents}
\label{Vents}
This simulation includes vents that account for the changes in ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). There are two ventilation scenarios that are considered in this study. The first is the baseline simulation where the ventilation times and ventilation areas represent our best understanding of the incident. The alternative simulation does not include the firefighter opening the rear door to make entry to the structure and the subsequent beta side wall breach. The times of the ventilation changes are provided in Table \ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of structure of this type. Therefore, leakage would have a negligible impact on the fluid mechanics within the structure and is not included in this study.

\bigskip
\begin{minipage}{.8\linewidth}
\centering
\captionof{table}{Simulation Ventilation Timelines}\label{tab:vents}
\begin{tabular}{ccccc}
\toprule[1.5pt]
Baseline Simulation & Alternative Simulation & Ventilation Area [m$^2$] & Type & Structure Side \\
\midrule
00    &  00   & 0.11 & Window      & Charlie \\
40    &  40   & 0.08 & Breach (x2) & Alpha and Charlie \\
45    &  45   & 0.12 & Breach      & Beta \\
55    &  55   & 0.12 & Breach      & Charlie \\     
130   &  N/A  & 1.41 & Door        & Charlie \\
135   &  N/A  & 0.2  & Breach      & Beta \\        
160   &  160  & 0.80 & Door        & Interior \\
300   &  300  & \multicolumn{3}{c}{Simulation Ends}  \\
\bottomrule[1.25pt]
\end{tabular}\par
\bigskip
\end{minipage}

@@

For the simulation in this study, we can estimate a measure of how well the flow field is resolved by the non-dimensional expression $D^*/\dx$. Here $D^*$ is a characteristic fire diameter, $\dx$ is the nominal size of a mesh cell, and $\dQ$ is the total heat release rate of the fire:
\begin{equation}
D^* = \left(
     \frac{\dQ}{\rho_\infty \, c_p \, T_\infty \, \sqrt{g} }
     \right)^\frac{2}{5} 
\label{eq:mesh}
\end{equation}   
From the FDS User Guide~\cite{FDS_Users_Guide}, ``the characteristic fire diameter is related to the characteristic fire size via the
relation $Q^* = (D^*/D)^{5/2}$.'' Here, $D$ is the physical diameter of the fire as input to the model. Following Eq.~(\ref{eq:mesh}) and using 10 cm grid cells, the characteristic fire diameter to cell size ratio based on the 3.3 MW porch fire is 15.5 and for the 9 MW fire is 23. Based on validation work performed for the U.S. Nuclear Regulatory Commission, $D^*/\dx$ values ranged between 4 and 1~\cite{NUREG_1824}. The grid resolution used in the model falls within typical, engineering $D^*/\dx$ values. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10 cm. Fig.~(\ref{fig:geom_grid}) shows the 
alpha side (front) of the structure rendered in Smokeview with the 10 cm mesh.

@@

As discussed Section~\ref{geom}, the structure has 17.15 m (56 ft) by 6.15 m (20ft) footprint with a total height of 8.56 m (28 ft). To ensure adequately resolved fluid flow into and out of the structure the computational domain is extended beyond the volume of the structure. The total volume of the computational domain is 24 m (79 ft) by 10 m (33 ft) by 10 m (33 ft). At 10 cm (3.9 in) resolution this would result in 2.4 million computational cells. To make the simulation time more tractable, the domain was divided into 16 equal 150 thousand cell meshes to be processed in parallel. Fig.~(\ref{fig:mult_mesh}) shows the entire structure within the computational domain.

\begin{figure}[h!]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_mesh}
\caption{Alpha-Delta side of the structure within computational domain. Each of the 16 meshes is 3 m (9.8 ft) by 10 m (33 ft) by 5 m (16.4 ft).}
\label{fig:mult_mesh}
\end{figure}

\chapter{Model Results}
\label{results}

@@

In Fig.~(\ref{fig:hrr}), the solid line is the baseline FDS simulation (this represents the on-scene conditions). The dashed line represents an alternative series events, specifically the rear door does not get opened and the subsequent beta side breach does not occur. The dash-dotted line is the prescribed FDS input HRR. We can see that within the first 10 seconds, there are under-ventilated conditions within the structure as the modeled HRR is below the prescribed HRR. The HRR peaks at just under 9 MW (approximately 20 seconds into the simulation) before becoming oxygen limited. The HRR drops to 5 MW as the available oxygen inside the structure gets consumed. There is insufficient available oxygen in the structure to combust all of the fuel and combustion occurs locally at vents. Fig.~(\ref{fig:smv_ext_fire}) illustrates exterior flaming combustion through the failed Charlie side window and attic breach just after the second fire ignites at 40s. After additional wall breaches at 45 and 55 seconds (see Table~\ref{tab:vents}), the abrupt change in ventilation causes the HRR spikes to 9 MW before again dropping to about 6 MW. The HRR then begins to increase towarda a sustainable steady state value.

From Fig.~(\ref{fig:hrr}), we see that the HRR for both FDS simulations are identical until 130 seconds (HRR $\approx$ 7 MW). At 130 seconds the rear door is opened in the baseline simulation. At 135 seconds there is an additional beta side breach (Table~\ref{tab:vents}). The open door and breach act as additional pressure releases for the structure. Again, the abrupt change in ventilation causes the HRR to spike. Here we see a rise from 7 MW to 9.7 MW at 139 seconds. At the time of door failure (160 seconds), the HRR in both cases was approximately 9 MW as there was sufficient ventilation in both cases. While there might be some locally different burning characteristics between the baseline simulation and the alternate simulation, in both cases the fire represented a significant hazard to firefighters on the alpha side of the second floor porch door.

@@

\begin{figure}[h!]
\includegraphics[width=.8\textwidth]{../Figures/west_50th_baseline_129}
\input{../Scripts/colorbar_metric} 
\caption{Temperature 1 second prior to rear door opening.}
\label{fig:temp_129s}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\textwidth]{../Figures/west_50th_baseline_159}
\input{../Scripts/colorbar_metric} 
\caption{Temperature 1 second prior to interior door failing.}
\label{fig:temp_159s}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.8\textwidth]{../Figures/west_50th_baseline_161}
\input{../Scripts/colorbar_metric} 
\caption{Temperature 1 second after interior door fails.}
\label{fig:temp_161s}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_175}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature 15 seconds after interior door fails.}
\label{fig:temp_175s}
\end{figure}
